An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-12-01 paper/1.introduction.tex

One direct solution to avoid this repeated computation of $get$ is to compute the composition in a right associative manner. For instance, if we transform  $lp3$ to $rp3$:
\[
rp3 = phead \circ (phead \circ phead)
\]
then the $put$ for $rp3$ only needs to compute $get$ of $phead$ twice, one time less than that for $lp3$.
 However, this transformation is not always easy to do. For instance, let us consider $breverse$, a bidirectional version of the traditional `reverse' program for reversing a list. It is defined using $\mbox{\it bfoldr}$, a bidirectional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}. Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
 \[
 \mbox{\it bfoldr}~bx~\cdots = \cdots  (\mbox{\it bfoldr}~\cdots) \circ bx \cdots
 \]
where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list. This make it hard to do the above transformation statically.

@@

In this paper, we made the first attempt for seriously considering the efficiency of evaluating BX compositions, and
solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.
The first method uses straightforward memoization: ``keeping \emph{intermediate states in a table} and using them when needed''. This avoids repeated $get$ computations and improves the runtime in many cases. However, this simple memoization needs to keep and search values in a table, which may introduce big cost for large inputs. We explain this method in Section~\ref{sect:minbigulm}.

>>>>

One direct solution to avoid this repeated computation of $get$ is to compute the composition in a right associative manner. For instance, if we transform  $lp3$ to $rp3$:
\[
rp3 = phead \circ (phead \circ phead)
\]
then the $put$ for $rp3$ only needs to compute $get$ of $phead$ twice, one time less than that for $lp3$.
 However, this transformation is not always easy to do. For instance, let us consider $breverse$, a bidirectional version of the traditional `reverse' program for reversing a list. It is defined using $\mbox{\it bfoldr}$, a bidirectional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}. Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
 \[
 \mbox{\it bfoldr}~bx~\cdots = \cdots  (\mbox{\it bfoldr}~\cdots) \circ bx \cdots
 \]
where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list. This makes it hard to do the above transformation statically.

@@

In this paper, we make the first attempt for seriously considering the efficiency of evaluating BX compositions, and
solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.
The first method uses straightforward memoization: ``keeping \emph{intermediate states in a table} and using them when needed''. This avoids repeated $get$ computations and improves the runtime in many cases. However, this simple memoization needs to keep and search values in a table, which may introduce big cost for large inputs. We explain this method in Section~\ref{sect:minbigulm}.

### 2019-12-01 paper/5.experiments.tex

We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default. There is no type system in our implementation\footnote{\url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}, so instead of constructing some dummies which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.

@@

We have conducted 7 test cases shown in Table~\ref{tab:test-cases}. Although the number of programs is not so much, they are not trivial and can be completely used in practice. In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table. These $bx$ are left associative composition since we we are only interested in this kind of program in the paper. Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

@@

The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from $n + 1$ similar \textit{pure} programs. We use the term \textit{pure} here to refer non-recursion compositions. The prefix \textit{lcomp} means that the associate precedence of $\circ$ in a straight-line composition is left. The suffix \textit{ldata} indicates that the input size is large. The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$. These structures are only intended to generate data enough large for observing experimental results.

>>>>

We have implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default. No type system in our implementation.}$^,$%HACK!
%
\footnote{Anonymous \url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}
%
described in the previous sections.
Instead of constructing some dummies, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.

@@

We have conducted 7 test cases shown in Table~\ref{tab:test-cases}. Although the number of cases is small, 
% they are not trivial and can be completely used in practice. 
they represent non-trivial cases of practical significance.
In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table. These $bx$ are left associative compositions since we we are only interested in this kind of program in the paper. Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

@@

% The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from the same $n + 1$ \textit{pure} programs. 
The first five test cases
% (no.~1--5) 
are straight-line compositions of the same $n + 1$ non-recursive programs. We refer to them as `pure' compositions.
% We use the term \textit{pure} here to refer non-recursion compositions. 
The prefix \textit{lcomp} in a test name indicates that the compositions are left-associative.
% that the associate precedence of $\circ$ in a straight-line composition is left.
The suffix \textit{ldata} indicates that the input size is large. The symbol $L$ in the input column of the \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$. These structures are only intended to generate test data that is large enough for 
% observing 
measuring
% experimental 
results.
